---
title: "EPI 560: Lab 07"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

if (!require("BiocManager", quietly = TRUE))
install.packages("BiocManager")
BiocManager::install("rhdf5")
#Read in files
library(readxl)
library(haven)
library(arrow)
library(hdf5r)

```


## Lab 7: Responses  

### Read in Files
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Question 1: Reading in files
## Question 2: Read times
# csv
before <- Sys.time()
pdemo_csv <- read.csv("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.csv")
after <- Sys.time()
csvtime <- as.numeric(after - before, units = "secs")

# Excel
before <- Sys.time()
pdemo_xlsx <- read_excel("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.xlsx")
after <- Sys.time()
exceltime <- as.numeric(after - before, units = "secs")

# Zip
before <- Sys.time()
#Call the first file in your zip folder csv_file
csv_file <- unzip("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.csv.zip", list = TRUE)$Name[1]
#Read in the file
pdemo_zip <- read.csv(unz("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.csv.zip", csv_file))
after <- Sys.time()
ziptime <- as.numeric(after - before, units = "secs")

# dta
before <- Sys.time()
pdemo_dta <- read_dta("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.dta")
after <- Sys.time()
dtatime <- as.numeric(after - before, units = "secs")

# feather
before <- Sys.time()
pdemo_feather <- read_feather("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.feather")
after <- Sys.time()
feathertime <- as.numeric(after - before, units = "secs")

# hdf5
before <- Sys.time()
#Explore the hdf5 file to see how many files there are
h5file <- H5File$new("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.hdf5", mode = "r")
#h5file$ls()
#Read in the file
pdemo_hdf5 <- h5file[["P_DEMO"]][]
after <- Sys.time()
hdf5time <- as.numeric(after - before, units = "secs")

# parquet
before <- Sys.time()
pdemo_parquet <- read_parquet("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.parquet")
after <- Sys.time()
parquettime<-after - before
parquettime <- as.numeric(after - before, units = "secs")


# R Data
before <- Sys.time()
load("C:/Users/stlp/OneDrive - UW/UW MPH/2025 Spring/EPI 560 - Data Management/R/data/P_DEMO.RData")
after <- Sys.time()
rdatatime <- as.numeric(after - before, units = "secs")

```


### Plot Files
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Question 1: Plot files
# csv
ggplot(pdemo_csv, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "red") +
  labs(
    title = "Scatterplot of P_DEMO csv: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()

# Excel
ggplot(pdemo_xlsx, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "orange") +
  labs(
    title = "Scatterplot of P_DEMO xlsx: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()

# Zip
ggplot(pdemo_zip, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "gold") +
  labs(
    title = "Scatterplot of P_DEMO zip: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()

# dta
ggplot(pdemo_dta, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "green") +
  labs(
    title = "Scatterplot of P_DEMO dta: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()

# feather
ggplot(pdemo_feather, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "blue") +
  labs(
    title = "Scatterplot of P_DEMO feather: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()

# hdf5
ggplot(pdemo_hdf5, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "cyan") +
  labs(
    title = "Scatterplot of P_DEMO hdf5: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()


# parquet
ggplot(pdemo_parquet, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "purple") +
  labs(
    title = "Scatterplot of P_DEMO parquet: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()


# R Data
ggplot(data_file, aes(x = WTINTPRP, y = WTMECPRP)) +
  geom_point(color = "magenta") +
  labs(
    title = "Scatterplot of P_DEMO RData: Full Sample Interview vs. MEC Exam Weight",
    x = "Interview Weight",
    y = "MEC Exam Weight"
  ) +
  theme_bw()


```

**1. All the data seem to be present in each format as each plot looks similar, if not identical. I noticed that loading the zip and hdf5 files was slightly more complex than the other file types since I had to retrieve the file before I could read it in.**

**2. The fastest file to load was the feather file which loaded in `r feathertime` seconds. This was `r parquettime-feathertime` seconds faster than it took the parquettime file to load. The slowest file to load was the Excel file which loaded in `r exceltime` seconds. This was `r exceltime-ziptime` seconds slower than the load time to load the Zip file.**

\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
