---
title: "EPI 560: Lab 06"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```


## Lab 6: Responses  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# code to set up SQL
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
#dbListTables(db)

```


### Part I: Simple query nesting

1. Write a query to generate a frequency table of reported hours of sleep among NHANES participants.
This table should have two columns: one named SleepHours, showing the number of hours of reported
sleep, and another named NumberReporting, showing the number of participants who reported the
corresponding number of hours of sleep.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 1

dbGetQuery(db, 
"WITH SelDat AS (
  SELECT 
    s.SEQN, 
    s.SLD012
      FROM P_SLQ s
  )
  SELECT
    COUNT(SEQN) AS N_Participants,
    SUM(SLD012) AS TotalSleep
      FROM SelDat"
)

```



2. Now modify that query so you return only the row where people are sleeping for 7.0 hours

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 2

dbGetQuery(db, 
"WITH SelDat AS (
  SELECT 
    s.SEQN, 
    s.SLD012
      FROM P_SLQ s
        WHERE SLD012 = 7.0
  )
  SELECT
    COUNT(SEQN) AS N_Participants,
    SUM(SLD012) AS Sleep
      FROM SelDat"
)

```

3. Now modify that query so you return only the modal row (i.e. the number the most people are reporting).
Identify that row using a query.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 3

##Identify the mode
dbGetQuery(db,
  "SELECT 
    s.SLD012 AS Mode
      FROM P_SLQ s 
        GROUP BY s.SLD012
        ORDER BY COUNT(*) DESC
        LIMIT 1")


##Query by the mode
dbGetQuery(db, 
"WITH SelMode AS (
  SELECT 
    s.SEQN, 
    s.SLD012
      FROM P_SLQ s
        WHERE SLD012 = (
  SELECT 
    s.SLD012 AS Mode
      FROM P_SLQ s 
        GROUP BY s.SLD012
        ORDER BY COUNT(*) DESC
          LIMIT 1
          )
  )
  SELECT
    COUNT(SEQN) AS N_Participants,
    SUM(SLD012) AS Sleep
      FROM SelMode"
)

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
dbDisconnect(db)
```

\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
