---
title: "EPI 560: Lab 05"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```



## Lab 5: Responses  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# code to set up SQL
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
#dbListTables(db)
query_auq <- dbGetQuery(db, "PRAGMA table_info('P_AUQ')")
query_slq <- dbGetQuery(db, "PRAGMA table_info('P_SLQ')")
query_demo <- dbGetQuery(db, "SELECT * FROM P_DEMO")

```



### Part I

1. Use SQL COUNT identify how many participants are included in the demographics table.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# code for Part I
# Question 1
dbGetQuery(db, "SELECT COUNT(*) AS DemoCount FROM P_DEMO")

```

**`r dbGetQuery(db, "SELECT COUNT(*) AS DemoCount FROM P_DEMO")$DemoCount` participants are included in the demographics table.**



2. Use SQL COUNT and INNER JOIN identify how many participants are included in both the demographics table and the sleep quality table.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 2
dbGetQuery(db, "SELECT COUNT(*) AS DemoCount FROM P_DEMO INNER JOIN P_SLQ ON P_DEMO.SEQN = P_SLQ.SEQN")

```

**`r dbGetQuery(db, "SELECT COUNT(*) AS DemoCount FROM P_DEMO INNER JOIN P_SLQ ON P_DEMO.SEQN = P_SLQ.SEQN")$DemoCount` participants are included in both the demographics and the sleep quality table.**



### Part II



3. Retrieve the first five lines (and only the first 5 lines) of your joined dataset using a SELECT query. Your final query should pull the variables that you need to answer the following questions, but should not contain variables that will not be needed.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 3
##Join all columns from different tables and print the first five lines
dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054, a.AUQ060, a.AUQ070, a.AUQ080,
  s.SLQ300, s.SLD012,
  d.RIAGENDR
FROM P_AUQ a
INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
INNER JOIN P_DEMO d ON a.SEQN = d.SEQN
LIMIT 5")

```



4. Among participants who have either good or excellent hearing, what is the average usual sleep time on weekdays or workdays that they get? What about among participants with a little trouble, moderate trouble, or a lot of trouble with their hearing?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
##Good or excellent hearing
avg_secs <- dbGetQuery(db, "
  SELECT 
    AVG(strftime('%s', '2000-01-01 ' || SLQ300)) AS avg_seconds
  FROM P_AUQ a
  INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
  INNER JOIN P_DEMO d ON a.SEQN = d.SEQN
  WHERE a.AUQ054 IN ('1', '2')
")$avg_seconds
###Convert avg_seconds back to HH:MM in R
avg_time <- format(as.POSIXct(avg_secs, origin = '1970-01-01', tz = 'UTC'), "%H:%M")


##A little, moderate, or a lot of trouble hearing
avg_secs <- dbGetQuery(db, "
  SELECT 
    AVG(strftime('%s', '2000-01-01 ' || SLQ300)) AS avg_seconds
  FROM P_AUQ a
  INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
  INNER JOIN P_DEMO d ON a.SEQN = d.SEQN
  WHERE a.AUQ054 IN ('3', '4', '5')
")$avg_seconds
###Convert avg_seconds back to HH:MM in R
avg_time_trouble <- format(as.POSIXct(avg_secs, origin = '1970-01-01', tz = 'UTC'), "%H:%M")

```

**Among participants who have either good or excellent hearing, the average usual sleep time on weekdays or workdays is `r avg_time`. Among participants with a little trouble, moderate trouble, or a lot of trouble with their hearing, the average is `r avg_time_trouble`.**



5. Write a new SQL query that returns the data from question 4 but also includes age of the included participants in years. Print the first six lines of the dataset that you retrieved.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 5
## Full dataset
dat5 <- dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054, a.AUQ060, a.AUQ070, a.AUQ080,
  s.SLQ300, s.SLD012,
  d.RIAGENDR, d.RIDAGEYR
FROM P_AUQ a
INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
INNER JOIN P_DEMO d ON a.SEQN = d.SEQN")

## Print first 6 lines
dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054, a.AUQ060, a.AUQ070, a.AUQ080,
  s.SLQ300, s.SLD012,
  d.RIAGENDR, d.RIDAGEYR
FROM P_AUQ a
INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
INNER JOIN P_DEMO d ON a.SEQN = d.SEQN
LIMIT 6")

```



6. Use R with the results from your prior query to determine many NHANES participants are missing values for the variables assessing whether individuals can hear a whisper, normal voice, or shout from across a quiet room. Is this missingness differential after stratifying at age 50 (note: we will consider age as a continuous variable later in the quarter)? Is it differential by usual weekday sleep time over 8 hours (note: we will consider usual weekday sleep time as a continuous variable later in the quarter)? How do you know?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 6
##Stratify
dat5$age_group <- ifelse(dat5$RIDAGEYR < 50, "<50", ">=50")
dat5$sleep_time <- ifelse(dat5$SLD012 < 8, "<8", ">=8")

##Whisper
overall_summary <- dat5 %>%
  summarise(
    total = n(),
    missing_whisper = sum(is.na(AUQ060))
  )

stratified_summary_age <- dat5 %>%
  group_by(age_group) %>%
  summarise(
    total = n(),
    missing_whisper = sum(is.na(AUQ060))
  )

stratified_summary_sleep <- dat5 %>%
  group_by(sleep_time) %>%
  summarise(
    total = n(),
    missing_whisper = sum(is.na(AUQ060))
  )

combined_summary_whisper <- bind_rows(overall_summary, stratified_summary_age, stratified_summary_sleep)

```

+ **The total number of participants missing values for the variable assessing hearing a whisper is `r combined_summary_whisper$missing_whisper[1]`. This missingness is differential after stratifying at age 50. Among participants under age 50, missingness for the variable assessing hearing a whisper is `r combined_summary_whisper$missing_whisper[2]` but among participants age 50 or older, missingness is `r combined_summary_whisper$missing_whisper[3]`. This missingness is also differential by usual weekday sleep time over 8 hours. Among participants who sleep under 8 hours during weekdays, missingness of the variable assessing hearing a whisper is `r combined_summary_whisper$missing_whisper[4]` but among those who sleep 8 or more hours, missingness is `r combined_summary_whisper$missing_whisper[5]`.**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
##Normal voice
overall_summary <- dat5 %>%
  summarise(
    total = n(),
    missing_normal = sum(is.na(AUQ070))
  )

stratified_summary_age <- dat5 %>%
  group_by(age_group) %>%
  summarise(
    total = n(),
    missing_normal = sum(is.na(AUQ070))
  )

stratified_summary_sleep <- dat5 %>%
  group_by(sleep_time) %>%
  summarise(
    total = n(),
    missing_normal = sum(is.na(AUQ070))
  )

combined_summary_normal <- bind_rows(overall_summary, stratified_summary_age, stratified_summary_sleep)

```

+ **The total number of participants missing values for the variable assessing hearing a normal voice is `r combined_summary_normal$missing_normal[1]`. This missingness is differential after stratifying at age 50. Among participants under age 50, missingness for the variable assessing hearing a normal voice is `r combined_summary_normal$missing_normal[2]` but among participants age 50 or older, missingness is `r combined_summary_normal$missing_normal[3]`. This missingness is also differential by usual weekday sleep time over 8 hours. Among participants who sleep under 8 hours during weekdays, missingness of the variable assessing hearing a normal voice is `r combined_summary_normal$missing_normal[4]` but among those who sleep 8 or more hours, missingness is `r combined_summary_normal$missing_normal[5]`.**


```{r, echo=FALSE, message=FALSE, warning=FALSE}
##Shout voice
overall_summary <- dat5 %>%
  summarise(
    total = n(),
    missing_shout = sum(is.na(AUQ080))
  )

stratified_summary_age <- dat5 %>%
  group_by(age_group) %>%
  summarise(
    total = n(),
    missing_shout = sum(is.na(AUQ080))
  )

stratified_summary_sleep <- dat5 %>%
  group_by(sleep_time) %>%
  summarise(
    total = n(),
    missing_shout = sum(is.na(AUQ080))
  )

combined_summary_shout <- bind_rows(overall_summary, stratified_summary_age, stratified_summary_sleep)

```

+ **The total number of participants missing values for the variable assessing hearing a shout is `r combined_summary_shout$missing_shout[1]`. This missingness is differential after stratifying at age 50. Among participants under age 50, missingness for the variable assessing hearing a shout is `r combined_summary_shout$missing_shout[2]` but among participants age 50 or older, missingness is `r combined_summary_shout$missing_shout[3]`. This missingness is also differential by usual weekday sleep time over 8 hours. Among participants who sleep under 8 hours during weekdays, missingness of the variable assessing hearing a shout is `r combined_summary_shout$missing_shout[4]` but among those who sleep 8 or more hours, missingness is `r combined_summary_shout$missing_shout[5]`.**


7. Write a SQL query that returns a dataframe with all participants who have a general condition of hearing that is either good or excellent and who have a usual sleep time on weekdays/workdays that is greater than 8. Print the first six lines of the dataset that you retrieved.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Question 7
## Print first 6 lines
dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054, a.AUQ060, a.AUQ070, a.AUQ080,
  s.SLQ300, s.SLD012,
  d.RIAGENDR, d.RIDAGEYR
FROM P_AUQ a
INNER JOIN P_SLQ s ON a.SEQN = s.SEQN
INNER JOIN P_DEMO d ON a.SEQN = d.SEQN
WHERE (a.AUQ054 = '1' OR a.AUQ054 = '2')
  AND s.SLD012 > 8
LIMIT 6")

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
dbDisconnect(db)
```

\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
