---
title: "EPI 560: Lab 04"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```


## Lab 4: Responses  


### Part I


2. The names of the 3 columns in P_HDL are SEQN, LBDHDD, LBDHDDSI


3. The highest value LBDHDD in the first three rows is 72.0

```{r part_1, echo=FALSE, message=FALSE, warning=FALSE}
# code for part 1
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
#Show all tables in db
#dbListTables(db)
query <- dbGetQuery(db, "SELECT * FROM P_HDL LIMIT 3")

```


### Part II

2. The columns and d-types are listed below under "name" and "type"

```{r part_2, echo=FALSE, message=FALSE, warning=FALSE}
# code for part 2
query <- dbGetQuery(db, "PRAGMA table_info('P_AUQ')")
query
```


3. The primary key for Audiometry (P_AUQ) is SEQN.


4. There are no foreign keys in this table.


5. This is a preview of the full P_AUQ table:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_auq <- dbGetQuery(db, "SELECT * FROM P_AUQ")
kable(head(p_auq, 5))
```


6. This is a preview of 2 columns, *SEQN* and *AUQ400 - When began to have hearing loss?*, from the P_AUQ table:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_auq_hearingloss <- dbGetQuery(db, "SELECT SEQN, AUQ400 FROM P_AUQ")
kable(head(p_auq_hearingloss, 5))
```


7. This is a preview of 2 columns, *SEQN* and *AUQ400 - When began to have hearing loss?*, filtering rows by *AUQ410b - Cause of hearing loss-Ear infections*, from the P_AUQ table:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_auq_hearingloss_infection <- dbGetQuery(db, "SELECT SEQN, AUQ400 FROM P_AUQ WHERE AUQ410b='2'")
kable(head(p_auq_hearingloss_infection, 5))
```


8. This is a preview of 2 columns, *SEQN* and *AUQ400 - When began to have hearing loss?*, filtering rows by *AUQ410b - Cause of hearing loss-Ear infections* and *AUQ430 - Ever had 3 or more ear infections?*, from the P_AUQ table:

```{r, echo=FALSE, message=FALSE, warning=FALSE}
p_auq_hearingloss_infection_freq <- dbGetQuery(db, "SELECT SEQN, AUQ400 FROM P_AUQ WHERE AUQ410b='2' AND AUQ430='1'")
kable(head(p_auq_hearingloss_infection_freq, 5))

```


### Part III

1. This is a table of glycohemoglobin (P_GHB) and audiometry (P_AUQ) joined by SEQN

*Help: How to remove the duplicate column for the primary ID?*

```{r part_3, echo=FALSE, message=FALSE, warning=FALSE}
# code for part 3
p_ghb_auq <- dbGetQuery(db, "SELECT * FROM P_GHB LEFT JOIN P_AUQ ON P_GHB.SEQN = P_AUQ.SEQN")
kable(head(p_ghb_auq, 5))
dbDisconnect(db)

```



/pagebreak  

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
