---
title: "EPI 560: Lab 11"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
library(lubridate)
library(summarytools)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```


## Lab 11: Responses 

### Part I: Planning and Documentation

**My plan is to look at whether having trouble hearing is associated with ever telling a doctor one had trouble sleeping, adjusted for age, and whether this association is modified by glycohemoglobin %.**

#### Question 1

a) SEQL tables and / b) variables:

+ Having trouble hearing (Categorical): P_AUQ / AUQ054

+ Telling a doctor one had trouble sleeping (Binary): P_SLQ / SLQ050

+ Age at screening (Continuous): P_DEMO / RIDAGEYR

+ Glycohemoglobin (Continuous): P_GHB/ LBXGH


c) Cleaning steps: 

1. rename variables

2. recategorize hearing 1-2 as "Good", 3-5 as "Trouble", 6 as "Deaf" 

3. recode 77, 99, 7, 9, "." as NA

4. relabel 1/2 as "Yes"/"No"

5. Drop participants who are "Deaf"


#### Question 2
Steps 1-4 will make reading the code and output easier. Step 5 may shift the cleaned data slightly from the raw data if there was an association between deafness and having trouble hearing.

#### Question 3
Steps 1-4 are things that I know I want to change. There is room for subjective decision making for step 5 in a sensitivity analysis.


### Part II: Data Cleaning

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Question 4
#Get data
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
dat <- dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054 AS Hearing,
  s.SLQ050 AS Sleep,
  g.LBXGH AS Glyco,
  d.RIDAGEYR AS Age
FROM P_AUQ a
LEFT JOIN P_SLQ s ON a.SEQN = s.SEQN
FULL JOIN P_GHB g ON a.SEQN = g.SEQN
FULL JOIN P_DEMO d ON a.SEQN = d.SEQN"
)
# Remove the duplicate SEQN column
dat <- dat[, !duplicated(names(dat))]
dbDisconnect(db)

#Clean data
##Raw data fixing missing
datcln <- dat %>%
  mutate(
    Hearing = ifelse(Hearing %in% c(77, 99, "."), NA, Hearing),
    
    Sleep = ifelse(Sleep %in% c(7, 9, "."), NA, Sleep),
    Sleep = case_when(
      Sleep == 1 ~ "Yes",
      Sleep == 2 ~ "No"
    ),
    
    Glyco = ifelse(Glyco %in% c("."), NA, Glyco),
    
    Age = ifelse(Age %in% c("."), NA, Age)
  )

##Cleaned data for Lab 11
datcln_lab11 <- dat %>%
  mutate(
    Hearing = ifelse(Hearing %in% c(77, 99, "."), NA, Hearing),
    Hearing = case_when(
      Hearing %in% 1:2 ~ "Good",
      Hearing %in% 3:5 ~ "Trouble",
      Hearing == 6 ~ "Deaf",
      TRUE ~ NA_character_
    ),
    
    Sleep = ifelse(Sleep %in% c(7, 9, "."), NA, Sleep),
    Sleep = case_when(
      Sleep == 1 ~ "Yes",
      Sleep == 2 ~ "No"
    ),
    
    Glyco = ifelse(Glyco %in% c("."), NA, Glyco),
    
    Age = ifelse(Age %in% c("."), NA, Age)
  ) %>%
  filter(Hearing != "Deaf")

```


#### Question 5

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Hearing Raw
barplot(table(datcln$Hearing),
     main = "Hearing Pre-Cleaning",
     ylab = "count",
     xlab = "Hearing Status")

## Hearing Cleaned
barplot(table(datcln_lab11$Hearing),
     main = "Hearing Post-Cleaning",
     ylab = "count",
     xlab = "Hearing Status")

```

**As I expected, it cleaning reduced relatively small counts from the "Deaf" group.**


#### Question 6
I will write my data out as a csv file so that it does not have to run again as a function. This is also more familiar to me than turning my data cleaning into a function. 


\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
