---
title: "EPI 560: Lab 12"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
library(lubridate)
library(summarytools)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```


## Lab 12: Responses 

### Part I: Get the Data

a) SEQL tables and / b) variables:

+ Age at screening (Continuous): P_DEMO / RIDAGEYR

+ Past year alcohol usage: P_ALQ / ALQ121

+ Telling a doctor one had trouble sleeping (Binary): P_SLQ / SLQ050

+ Having trouble hearing (Categorical): P_AUQ / AUQ054

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Part I
#Get data
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
dat <- dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054 AS HearingTrouble,
  s.SLQ050 AS Sleep,
  g.ALQ121 AS Alcohol,
  g.ALQ111 AS AlcoholEver,
  g.ALQ151 AS AlcoholBinge,
  d.RIDAGEYR AS Age
FROM P_AUQ a
LEFT JOIN P_SLQ s ON a.SEQN = s.SEQN
FULL JOIN P_ALQ g ON a.SEQN = g.SEQN
FULL JOIN P_DEMO d ON a.SEQN = d.SEQN"
)
# Remove the duplicate SEQN column
dat <- dat[, !duplicated(names(dat))]
dbDisconnect(db)

##Cleaning
datcln <- dat %>%
  mutate(
    HearingTrouble = ifelse(HearingTrouble %in% c(77, 99, "."), NA, HearingTrouble),
    HearingTrouble = case_when(
      HearingTrouble %in% 3:6 ~ "Yes",
      HearingTrouble %in% 1:2 ~ "No",
      TRUE ~ NA_character_
    ),
    
    Sleep = ifelse(Sleep %in% c(7, 9, "."), NA, Sleep),
    Sleep = case_when(
      Sleep == 2 ~ 0,
      Sleep == 1 ~ 1,
      TRUE ~ NA_real_
    ),
    
    Alcohol = ifelse(Alcohol %in% c(77, 99, "."), NA, Alcohol),
    
    AlcoholEver = ifelse(AlcoholEver %in% c(7, 9, "."), NA, AlcoholEver),
    
    AlcoholBinge = ifelse(AlcoholBinge %in% c(7, 9, "."), NA, AlcoholBinge),
    
    Age = ifelse(Age %in% c("."), NA, Age)
  )

```


## Part II: Fit a Model

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Code for Part II
options(scipen = 999)

##Factor alcohol usage
datcln$Alcohol <- as.factor(datcln$Alcohol)

##Poisson Regression
mod <- glm(Sleep ~ HearingTrouble + Age + Alcohol, family = poisson, data = datcln)
summary(mod)

#Relative risks
exp(coef(mod))
#Confidence interval
exp(confint(mod))

```

## Part III: Functionalize your model

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Code for Part III
##Step 1
rrci <- function(formula, dataset) {
  print("called my function")
}
rrci(formula='Sleep ~ HearingTrouble + Age + Alcohol', dataset=datcln)


##Step 2
rrci <- function(formula, dataset) {
  print(sprintf("called rrci with %s", deparse(formula)))
}
rrci(formula = Sleep ~ HearingTrouble + Age + Alcohol, dataset = datcln)


##Step 3
rrci <- function(formula, dataset) {
  print(sprintf("called rrci with %s", deparse(formula)))
  mod <- glm(formula, datcln, family=poisson)
  print("fit model successfully")
}
rrci(formula = Sleep ~ HearingTrouble + Age + Alcohol, dataset = datcln)


##Complete Function
rrci <- function(formula, dataset) {

  mod <- glm(formula, dataset, family=poisson)
  
  tidy_mod <- broom::tidy(mod, conf.int = TRUE, exponentiate = TRUE)
  
  main_var <- attr(terms(formula), "term.labels")[1]
  main_term <- tidy_mod[grepl(paste0("^", main_var), tidy_mod$term), ]

  return(list(
    relative_risk = main_term$estimate[1],
    ci_95 = c(main_term$conf.low[1], main_term$conf.high[1])
  ))
}

rrci(formula = Sleep ~ HearingTrouble + Age + Alcohol, dataset = datcln)

```

## Part IV: Consider different categorizations for alcohol use and age

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Code for Part IV
##1) Lifetime Ever Drinkers/Never Drinkers
datcln1 <- datcln %>%
  mutate(
    Alcohol = case_when(
      AlcoholEver == 1 ~ 1,
      AlcoholEver == 2 ~ 0,
      TRUE ~ NA_real_
    ))

##2) Past-year Ever Drinkers/Past-year Never Drinkers
datcln2 <- datcln %>%
  mutate(
    Alcohol = case_when(
      Alcohol != 0 ~ 1,
      Alcohol == 0 ~ 0,
      TRUE ~ NA_real_
    ))

##3) Past-year Ever-binge Drinkers/Never-binge Drinkers
datcln3 <- datcln %>%
  mutate(
    Alcohol = case_when(
      AlcoholBinge == 1 ~ 1,
      AlcoholBinge == 2 ~ 0,
      TRUE ~ NA_real_
    ))

##4) Past-year Never Drinkers/Daily Drinkers/In-between Drinkers
datcln4 <- datcln %>%
  mutate(
    Alcohol = case_when(
      #Never
      Alcohol == 0 ~ 0,
      #In-between
      Alcohol %in% 2:10 ~ 1,
      #Daily
      Alcohol == 1 ~ 2,
      TRUE ~ NA_real_
    ))

```

## Part V: Use the Function

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Code for Part V
#Estimate RR in original dataset
rrci(formula = Sleep ~ HearingTrouble + Age + Alcohol, dataset = datcln)

#Estimate RR in the first (lifetime ever/never drinkers) dataset
rrci(formula = Sleep ~ HearingTrouble + Age + Alcohol, dataset = datcln1)

```

## Part VI: Loop over all datasets

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#Code for Part VI
datasets <- list(
  "1) Original Dataset" = datcln,
  "2) Lifetime Ever Drinkers/Never Drinkers" = datcln1,
  "3) Past-year Ever Drinkers/Never Drinkers" = datcln2,
  "4) Past-year Ever-binge Drinkers/Never-binge Drinkers" = datcln3,
  "5) Past-year Never Drinkers/Daily Drinkers/In-between Drinkers" = datcln4
)

rr_values <- c()

for (name in names(datasets)) {
  result <- rrci(Sleep ~ HearingTrouble + Age + Alcohol, dataset = datasets[[name]])
  rr_values[name] <- result$relative_risk
}

rr_table <- data.frame(
  labels = names(rr_values),
  relative_risks = unname(rr_values)
)

print(rr_table)

```


\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
