---
title: "EPI 560: Lab 10"
author: "Gina Huynh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document:
        df_print: paged
        theme: cosmo
        toc: yes
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "build") })
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Setup
library(pacman)
library(DBI)
library(RSQLite)
library(lubridate)
library(summarytools)
p_load("tidyverse", "knitr", "usethis", "gitcreds")

```


## Lab 10: Responses 

### Part I:

**My plan is to look at whether having trouble hearing is associated with ever telling a doctor one had trouble sleeping, adjusted for age, and whether this association is modified by glycohemoglobin %.**

+ Having trouble hearing (Categorical): P_AUQ / AUQ054

+ Telling a doctor one had trouble sleeping (Binary): P_SLQ / SLQ050

+ Age at screening (Continuous): P_DEMO / RIDAGEYR

+ Glycohemoglobin (Continuous): P_GHB/ LBXGH


#### Retrieve the data you want using RSQLite - Note that you will need the same data in R for future
assignments, so write your query so you can re-use it. How many rows and columns do you have?

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Part I
#Get data
db <- dbConnect(RSQLite::SQLite(), dbname = "epi560.sqlite")
dat <- dbGetQuery(db, "SELECT 
  a.SEQN, 
  a.AUQ054 AS Hearing,
  s.SLQ050 AS Sleep,
  g.LBXGH AS Glyco,
  d.RIDAGEYR AS Age
FROM P_AUQ a
LEFT JOIN P_SLQ s ON a.SEQN = s.SEQN
FULL JOIN P_GHB g ON a.SEQN = g.SEQN
FULL JOIN P_DEMO d ON a.SEQN = d.SEQN"
)
# Remove the duplicate SEQN column
dat <- dat[, !duplicated(names(dat))]
dbDisconnect(db)

#Clean data
datcln <- dat %>%
  mutate(
    Hearing = ifelse(Hearing %in% c(77, 99, "."), NA, Hearing),
    Hearing = case_when(
      Hearing %in% 1:2 ~ "Good",
      Hearing %in% 3:5 ~ "Trouble",
      Hearing == 6 ~ "Deaf",
      TRUE ~ NA_character_
    ),
    
    Sleep = ifelse(Sleep %in% c(7, 9, "."), NA, Sleep),
    Sleep = case_when(
      Sleep == 1 ~ "Yes",
      Sleep == 2 ~ "No"
    ),
    
    Glyco = ifelse(Glyco %in% c("."), NA, Glyco),
    
    Age = ifelse(Age %in% c("."), NA, Age)
  )


#Number of rows and columns
dim(dat)

```
**I have `r dim(dat)[1]` rows and `r dim(dat)[2]` columns in my dataset.**


### Part II: Get a sense of the distributional statistics of the data

#### (a) Make histograms or scatterplots of at least 2 numeric columns that you think will be useful.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Part IIa
hist(datcln$Age,
     main = "Age")

hist(datcln$Glyco,
     main = "Glycohemoglobin")

```


#### (b) Make bar plots for at least 2 categorical columns that you think will be useful.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Part IIb
barplot(table(datcln$Hearing),
     main = "Hearing",
     ylab = "count",
     xlab = "Hearing Status")

barplot(table(datcln$Sleep),
     main = "Trouble Sleeping",
     ylab = "count",
     xlab = "Trouble Sleeping")

```


#### (c) Make a table with summary statistics for each variable, including missingness
```{r, echo=FALSE, message=FALSE, warning=FALSE}
## Code for Part IIc

library(summarytools)
print(dfSummary(datcln[ , !(names(datcln) %in% "SEQN")]))

#library(skimr)
#skim(datcln)

```


\newpage

## Code Appendix  
 
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE} 
```
